Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative analyses

### 2018-03-01 paper-webserver.tex
EDIT: readthrough

Public health laboratories are currently moving to whole-genome sequencing based analyses, and require rapid prediction of relevent pathogen phenotypes impacting health. Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
\mwcomment{Reference laboratories do not care about data retrieval, only real-time results. Researchers care more about context and thus would use a historical record of phenotypes for all predicted strains} 
To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows. The newly developed Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative anlaysis of tens of thousands of genome sequences. Though generally applicable to bacterial genome sequences, Spfy currently contains X \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
Spfy links the results and metadata to the genome sequences through a standardized ontology, which facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the recomputing of analyses.
The graph approach is flexible, and can accommodate new analysis software modules as they are developed, and easily link new results to those already stored.
Integrated data storage and analyses are currently necessary as the number of publicly available whole genome sequences is currently in the hundreds of thousands, with millions likely to be available within the next few years. 
\mwcomment{should we mention how many ecoli genomes are available? It might prompt reviewers to ask why we don't have them in spfy}
\par

@@

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the identification of known antimircrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}. For clinical diagnoses and comparisons, individual species can be first divided into subtypes with complementing AMR and VF results. Software methods for subtyping rely on intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}. These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.

Comprehensive platforms that combine individual programs into a cohesive whole also exist. These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2016improvements}. Commercial applications, such as Bionumerics, which is used by PulseNet International for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}. These platforms are designed to be applied to individual projects \cite{}.

% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
WGS of bacterial pathogens have recently accumulated in public databases in the hundreds of thousands, with millions set to be available within the next few years. For \textit{Escherichia coli} alone, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million sequenced genomes in GenBank \cite{doi:10.1093/nar/gks1195}. Many of the comparative analyses that are currently used in the analyses of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes. An effective method to mitigate the recomputing of analyses, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology. Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
\mwcomment{should we mention how many ecoli genomes are available? It might prompt reviewers to ask why we don't have them in spfy}

@@

% 5. how we solved it
In this study, we present the Spfy update to the SuperPhy platform, which integrates a graph database with real-time analyses; this integration avoids recomputing identical analyses. Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed. The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

@@

Spfy provides rapid in-silco versions of common reference laboratory tests for the analyses of \textit{E. coli}. It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification {ectyper ref}; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli\_serotyping}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.

Spfy also performs pangenome analyses using Panseq \cite{laing2010pan}, and provides machine learning modules for biomarker discovery among groups using Scikit-learn \cite{pedregosa2011scikit}.

@@

By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological association studies. Any data type or relation in the graph is a valid option for analysis. This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source.

@@

i) The upload begins through the website, where user-defined analyses options are selected. The results of these analyses are immediately reported to the user following their completion, while all other non-selected analyses are subsequently completed in the background and stored in the database without interaction from the user. The public web service accepts uploads of up to 200 MB (approximately 50 \textit {E. coli} genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted to a local instance.

@@

iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Superphy.

\subsection{Data Storage}
% para
% 0. Goals: big-data, everything linked, easy addition of new links
% 1. spfy is built around graph technologies
% 1. how we structure our graph
% 2. ontoogies used
% 3. inferencing
% 3 1/2. SPARQL queries
% para: the semantic web
Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}. For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from given ontologies. This system is flexible and allows novel data to be incorporated into the existing graph.
\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/ontology}
\end{center}
\caption{Structure of Spfy's graph database. Brackets highlight the source of different data points and the software it was generated from. These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database. Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
\label{fig-ontology}
\end{figure}

@@

\section{RESULTS}
Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules. This included: serotypeing (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of pan-genome content \textit{E. coli}.
The resulting database had 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
Spfy has been up since May 2017. The server accepts assembled \textit{E. coli} genomes with the \textit{.fasta} or \textit{.fna} extensions. Submissions are subjected to quality control, ensuring the submitted genomes are \textit{E. coli} sequences before subsequent analyses are run.
\par

@@

% drawbacks - big data
Many bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
Recent efforts \cite{goecks2010galaxy,thomsen2016bacterial} have focused on providing a common web interface for these programs, while still returning the same result files.
However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required for types of analyses that need to be performed \cite{schatz2015biological}.

One of the key goals in developing Spfy was to accommodate and store a variety of result formats, and then to make the data from these results retrievable and usable as inputs for downstream analyses, such as predictive biomarker discovery. We have shown how a graph database can accommodate the results given by a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.

\mwcomment{I think this discussion sets the perfect tone for the big data applications provided by spfy. My only suggestion is to give examples of the types analysis that are mentioned in the review. One idea, instead of dropping "Impact on Public Heath Efforts" entirely, you can convert it into a use case example that highlights the benefit spfy provides.}

@@

% para
% focus on application
The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays to identify AMR, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}. Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory test.

Spfy provides rapid and easy predictive genomic analyses of \textit{E. coli} genomes while also addressing the problem of large scale comparisons. With the larger datasets involved in population genomics, it is no longer viable for individual researchers to download data to perform comparisons. Instead, efforts have focused on storing biological data online and enabling analyses on that data \cite{schatz2015biological}. By using a graph database, Spfy can integrate results from  different technologies including software predicted phenotypes as well as laboratory results and user-submitted metadata. In addition, datasets can be built and submitted from multiple labs for joint analysis.

@@

% namely galaxy
The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}. Galaxy aims to provide a reproducible, computational interface which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated and also extended as new analysis tools are developed. The Galaxy worklow focus is on running an individual analysis pipeline. It does not include functionality to store and collate analysis results for large-scale comparative analysis.

@@

Spfy is similar to these technologies in that it automates workflows for users, and as in Galaxy, uses task queues to distribute selected analyses.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy processes XXXX files over XX tasks in XXXX time, distributing computations over a task queue and multiple Docker compartmentalized containers.
However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for downstream comparative analysis. This allows Spfy to perform population-wide analyses, regardless of the individual analysis software used to generate the results for an individual genome.

@@

PATRIC \cite{wattam2016improvements} adds support for comparing up to nine user-submitted genomes against a reference genome, targeting various gene annotations. PATRIC indexes a NoSQL document store to compare similar document types and is beneficial for comparative genomics. Unlike PARTIC, Spfy is targeted towards statistical comparisons of any population grouping, and Spfy's graph database has no limit on the number of genomes grouped for comparison. For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in \ref{fig-groupings}.

@@

\mwcomment{Maybe should include a comparison to PATRIC. Patric does offer some comparative genomic analysis, however they are not health focused (pathways, gene presence/absence etc), and you cannnot add your own data to the repository}

@@

Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}. While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy. The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license. A developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.

>>>>

Public health laboratories are currently moving to whole-genome sequencing based analyses, and require rapid prediction of relevant pathogen phenotypes impacting health. Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
\mwcomment{Reference laboratories do not care about data retrieval, only real-time results. Researchers care more about context and thus would use a historical record of phenotypes for all predicted strains} 
To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows. The newly developed Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative analysis of tens of thousands of genome sequences. Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
Spfy links the results and metadata to the genome sequences through a standardized ontology, which facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the recomputing of analyses.
The graph approach is flexible, and can accommodate new analysis software modules as they are developed, and easily link new results to those already stored.
Integrated data storage and analyses are currently necessary to match the rapid accumulation of whole genome sequences in public databases.
\par

@@

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}. For clinical diagnoses and comparisons, individual species can be first divided into subtypes with complementing AMR and VF results. Software methods for subtyping rely on pre-selected intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}. These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.

Comprehensive platforms that combine individual programs into a cohesive whole also exist. These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2016improvements}. Commercial applications, such as Bionumerics, which is used by PulseNet International for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}. These platforms are designed to be applied to individual projects.

Many of the comparative analyses that are currently used in the study of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes. An effective method to mitigate the recomputing of analyses, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology. Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.

@@

% 5. how we solved it
In this article, we present a new platform merging Superphy's pre-computed results with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical runs. Graph-based result storage also allows retrospective comparisons across stored results as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed. The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

@@

Spfy provides rapid in-silco versions of common reference laboratory tests for the analyses of \textit{E. coli}. It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}. Spfy also performs pangenome analyses using Panseq \cite{laing2010pan}, with the entire pangenome stored in the database and associated with source genomes.

@@

By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological association studies. Any data type or relation in the graph is a valid option for analysis. This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source. All results are displayed to users in real-time, usually within 2-3 minutes.

@@

i) The upload begins through the website, where user-defined analysis options are selected. The results of these analyses are immediately reported to the user following their completion, while all other non-selected analyses are subsequently completed in the background and stored in the database without interaction from the user. The public web service accepts uploads of up to 200 MB (approximately 50 \textit {E. coli} genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted to a local instance.

@@

iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Spfy.

\subsection{Data Storage}
% para
% 0. Goals: big-data, everything linked, easy addition of new links
% 1. spfy is built around graph technologies
% 1. how we structure our graph
% 2. ontoogies used
% 3. inferencing
% 3 1/2. SPARQL queries
% para: the semantic web
Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}. For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies. This system is flexible and allows novel data to be incorporated into the existing graph.
\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/ontology}
\end{center}
\caption{Structure of Spfy's graph database. Brackets highlight the source of different data points and the software it was generated from. These parts are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database. Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
\label{fig-ontology}
\end{figure}

@@

\section{RESULTS}
Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules. This included: serotyping (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of the pan-genome content of \textit{E. coli}.
The resulting database has 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
Spfy has been up since May 2017. The server accepts assembled \textit{E. coli} genomes with the \textit{.fasta} or \textit{.fna} extensions. Submissions are subjected to quality control, ensuring the submitted genomes are \textit{E. coli} sequences before subsequent analyses are run.
\par

@@

% drawbacks - big data
Many bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
Recent efforts \cite{goecks2010galaxy,thomsen2016bacterial} have focused on providing a common web interface for these programs, while still returning the same result files.
However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required for the types of analyses that need to be performed \cite{schatz2015biological}.

One of the key goals in developing Spfy was to accommodate and store a variety of result formats, and then to make the data from these results retrievable and usable as inputs for downstream analyses, such as predictive biomarker discovery. We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.

% \mwcomment{I think this discussion sets the perfect tone for the big data applications provided by spfy. My only suggestion is to give examples of the types analysis that are mentioned in the review. One idea, instead of dropping "Impact on Public Heath Efforts" entirely, you can convert it into a use case example that highlights the benefit spfy provides.}

@@

% para
% focus on application
The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays to identify AMR, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}. Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory tests.

Spfy provides rapid and easy predictive genomic analyses of \textit{E. coli} genomes while also addressing the problem of large scale comparisons. With the larger datasets involved in population genomics, it is no longer viable for individual researchers to download data to perform comparisons. Instead, efforts have focused on storing biological data online and enabling analyses on that data \cite{schatz2015biological}. By using a graph database, Spfy can integrate results from  different technologies including software predicted phenotypes as well as laboratory results and user-submitted metadata. In addition, datasets can be built and submitted from multiple labs for joint analyses.

@@

% namely galaxy
The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}. Galaxy aims to provide a reproducible, computational interface which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so the entire pipeline can be replicated and also extended as new tools are developed. The Galaxy workflow focuses on running an individual analysis pipeline. It does not include functionality to store and collate analysis results for large-scale comparative studies.

@@

Spfy is similar to these technologies in that it automates workflows for users, and as in Galaxy, uses task queues to distribute selected analyses.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, usually completing within 2-3 minutes, though Spfy does not provide genome assembly services.
However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for downstream comparative studies. This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.

@@

PATRIC \cite{wattam2016improvements} and Spfy share the same goal of integrated analyses. PATRIC adds support for comparing up to nine user-submitted genomes against a reference genome, targeting various gene annotations. The platform indexes a NoSQL document store to compare similar document types and is beneficial for comparative genomics. Unlike PARTIC, Spfy is targeted towards statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison. For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in \ref{fig-groupings}.

@@

% \mwcomment{Maybe should include a comparison to PATRIC. Patric does offer some comparative genomic analysis, however they are not health focused (pathways, gene presence/absence etc), and you cannnot add your own data to the repository}

@@

Future work will focus on adding additional analysis modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}. While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy. The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license. A developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
